Drivers of tree regeneration remain challenging to be captured by models of

### 2022-12-02 manuscript/manuscriptRW.Rmd
additions on the manuscript based on feedback from Harald on 24th November

---
title: Drivers of tree regeneration remain challenging to be captured by models of
  forest dynamics
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Mats Mahnken,
- Samuel Fischer,
- Xavier Morin,
- Manfred Lexer,
- Christian Hochauer,
- Katarina Merganicova,
- Tobias Mette,
- Werner Rammer,
- Kristin H. Braziunas,
- Paola Mairota,
- Josef Bruna,
- Heike Lischke,
- Daniel Scherrer,
- Simon Scheiter,
- Thomas Hickler,
- Tim Anders,
- Jessica Hetzer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 15 models and variants of forest
dynamics at the stand, landscape and global levels, we simulate the
regeneration across 200 sites representing environmental gradients
across Europe. The simulation results are compared with observed data
collected from forest reserves across Europe. In this study we want to
asses how well are models of forest dynamics capturing regeneration
levels, species composition and mortality in tree recruitment. The
results show that most of the models of forest dynamics tend to
over-estimate recruitment rates although this is not always connected to
large mortality rates in early dynamics stages. The effect of light
availability on recruitment rates is better captured than temperature or
soil moisture, however the patterns are not consistent across models.
Species richness matches quite well compared to the ranges observed in
the data. Models simulating higher species diversity at the stand level
do not feature higher diversity in recruitment. *Abies alba* was well
captured across temperature and soil moisture gradients by most models
and *Fagus sylvatica* was generally underestimated. *Picea abies*
recruitment was generally over-estimated on colder sites. HERE TWO
SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE

@@

Forests provide multiple products and services of vital importance.
Under the pressure of climate change and increasing disturbances, it is
even more important that we understand how forests will change and
evaluate their capacity to adapt or be resilient to these changes
[@lindner2010]. A wide range of models of forest dynamics were developed
over the past decades considering the impacts of climate in forest
dynamics [@bugmann2022]. From these studies, it is becoming increasingly
evident that while we have a reasonably good understanding how to model
tree growth [@vanclay1997; @bugmann1996], and substantial efforts have
been dedicated to improving the depiction of tree mortality processes
[@bugmann2019], tree regeneration remains a less studied process in most
of these models, while becoming ever more important in the context of
forest disturbances and forest resilience.

Tree recruitment depends on multiple processes such as reproduction,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival; these processes are all difficult to assess and
some of them are scarcely understood and thus appear highly stochastic
[@bogdziewicz2021]. Finding the right level of complexity in the
definition of the factors leading to tree recruitment is challenging and
models of forest dynamics have used different approaches [@bugmann2022].
Overall, models are needed to (1) synthesize existing empirical data and
explore their consequences, and (2) assess future tree recruitment e.g.
in the context of global change scenarios. Currently, regeneration
processes in dynamic forest models are handled in a multitude of ways
[@konig_tree_2022; @bugmann2022], from entirely ignoring tree
regeneration (as done in classical forest growth models, e.g.
@pretzsch2002), across the use of a few simple environmental filters, as
done in most forest "gap" models (e.g., @shuman2011) and Dynamic Global
Vegetation Models (e.g., @smith2001; @hickler2012), to complex
approaches that incorporate local feedback from the canopy, multiple
ecological processes and often also small time steps (e.g., @wehrli2006;
@seidl2012), or field-based statistical parameterizations, which however
are not easy to extrapolate in space and time (e.g., @ribbens1994).

With the current strategies that models of forest dynamics are using to
represent tree recruitment we are obtaining results that faced several
problems and have strong impacts on forest dynamic such as
over-recruitment, needed excess of mortality at early stages and
inadecuate species composition in regeneration. Some model outputs are
over-recruiting which subsequently must die primarily by competition in
later stages. This is characteristic of 'classical' forest gap models
[@gutiérrez2016], at least as long as the simulations comprise a
multitude of species. Other models use a model output to correct
recruitment levels for example by routinely calibrate them against the
predicted canopy species composition affecting as well the species
dominance.

Correctly capturing the species composition in recruitment is often also
a challenge, thus leading to unrealistic successional drift in the model
REFERENCE?, for example by the change of the dominant species via
stochastic processes that continuously bring seeds of other species that
the come to dominance, with this strategy a new species might start to
rise eliminating the former dominant species. A related issues is the
rarification of species richness due to positive feedback effects such
that eventually just single-species forest stands remain. This is
sometimes handled by the incorporation of a low level of seed influx of
all species at all times (e.g., @lischke2006 ). However, simulated
species composition is usually exceedingly sensitive to that level
whereas the parameters of such functions are poorly constrained by field
data. There is also often a problem with the insuficient constraint of
parameter values for models that start from very small tree sizes (e.g.,
10 cm of height or even seeds) and track tree development in a
process-oriented manner by considering a multitude of ecological
influences, rather than being statistically driven and emphasizing
recruitment into a larger size class (e.g., @wehrli2006 ). Most of these
undesired behaviors in tree recruitment are often not published, but in
the vast majority of studies they are corrected (e.g., @rasche2011,
where for most study sites regeneration was simply switched off). Thus,
a strong focus on the modeling of tree regeneration processes is highly
important and long overdue [@price] if we are to make reliable
projections of future forest dynamics, i.e. when the models need to be
operated in extrapolation mode, but also from a fundamental ecological
point of view (increasing systems understanding).

Tree recruitment is understudied in field studies and modelling
[@hanburybrown_forest_2022], as little data has been available and it is
often very fragmented, i.e., not in one harmonized data set. On the one
hand, monitoring on permanent sample plots (such as national forest
inventories) often measure tree recruitment [@zell_modeling_2019]
however often with a design that captures the regeneration intensity
after the stand initiation phase on productive forests [@hallsby2015].
Data from permanent plots, although available across regions, it is also
highly heterogeneous in design making their usability more difficult
[@käber2021]. On the other hand, targeted studies to measure tree
recruitment on experimental sites are often limited in space, species or
the relevant processes leading to regeneration that are targeted
[@berdanier2016; @collet2006].

This is the first time, a large number of forest models commonly used to
assess forest dynamics under climate change are evaluated against a
continental-scale multi-species harmonized data set on tree recruitment
[@käber2022]. In this study we relied on models with various
'philosophies', including models operating at the stand to the global
scale. Because of the large variability in recruitment patterns in
nature and the large amount of factor driving recruitment, including
some which are not included in most models, such as deer browsing, we do
not aim at a detailed statistical evaluation of each model. Instead, we
aim at evaluating general recruitment patterns, magnitudes and benchmark
the regeneration niche of multiple species based on empirical data along
a wide environmental gradient (temperature, moisture, and light
availability). More specifically in this study we wanted to answer the
following questions: 1) are models of forests dynamics capturing
correctly regeneration levels across environmental gradients; 2) are
models of forest dynamics representing correctly initial composition of
tree populations;and, 3) can models of forest dynamics correctly capture
the initial mortality in tree in-growth?.

@@

Fifteen models of forest dynamics (including two model variants) were
used to assess tree regeneration. The models selected simulate forest
dynamics at the stand, landscape and global scale (Table
\@ref(tab:tableModels)). Models construction philosophies and origin
differs, most of the models have a largely process based approach and
two models have formulations derived from statistical analyses based on
inventory data (SYBILA and xComp). Some of the models largely rely on
the original approach underlying forest gap models (ForClim 1, ForClim
11, LandClim, Landis II ); and some include plant ecophysiology
formulations (FORMIND, PICUS, 4C, iLand). This different approaches in
the models construction also has influenced how tree establishment is
defined.

Based on their establishment modules the models can be categorized as
regeneration or recruitment (or ingrowth) models [@vanclay1994;
@konig_tree_2022] (see Table \@ref(tab:tableModels)). Regeneration
models include processes such as flowering and pollination, seed
production, seed dispersal, germination or seedling establishment that
lead to the number of seedlings or saplings. Recruitment models
introduce a number of seedlings or saplings with certain characteristic
such as biomass or dbh, without considering earlier development
processes. We can also distinguish models that have feedback in the
regeneration as those in which the level and species composition in
regeneration is influenced by the existing adult tree community via the
production of seeds, seedlings or saplings. The starting point for
regeneration it is also different per model varying from seedbank, seed,
sapling or ingrowth. The overall complexity in the establishment modules
also varies, following @bugmann2022 we can see that most of the models
included in this study can be classified as having rather low complexity
in their tree recruitment modules while few of them have high complexity
specifically when defining processes such as XXXX, XXXX (enumerate
models).

Two models, ForClim and ForCEEPs used two alternative model versions.
HERE **FORCLIM** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCLIM 1
AND 11 AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS. HERE
**FORCEEPS** WRITE 2 SENTENCES ON THE DIFFERENCES BETWEEN FORCEEPS AND
FORCEEPS(F) AND WHY IT IS RELEVANT TO CONSIDER THE TWO MODEL VERSIONS.

@@

Regeneration data covering a wide range of environmental gradients is
hard to obtain and this is one of the reasons why most of the models of
forest dynamics have never been confronted with a data-set covering a
wide range of climate gradients in a large number sites to evaluate how
well are they capturing regeneration. The observed regeneration data
used in this study represent the available range of environmental
gradients in temperature and precipitation in Europe using 200 sites of
strict forest reserves compiled in the framework of the EuFoRIa network
(Figure \@ref(fig:map)). The sample forests have not been managed in at
least 10 years (most of them in close to 100 years?), and each of the
sites contains at least two consecutive measurements. Regeneration was
measured at 7 and 10 cm threshold in all the sites except for 35 sites
where only regeneration at 10 cm diameter threshold was available. For
more details on this dataset and their regeneration information please
refer to @käber2022.

@@

Each modeler group was provided with a detailed protocol (available in
the supplementary material) with instruction on how to perform the
simulations regarding input variables such as climate or soil,
simulation setup and expected output variables. No information was
provided on the regeneration prior to the simulation. The overarching
goal of this experiment was the assessment of the regeneration niche as
it arises from empirical data analysis against its representation in a
wide range of models of forest dynamics.

The input variables were collected and aggregated to be adapted to the
different model demands. The time series of climatic variables from 1981
to 2018 were provided in hourly (Era5-land data, @muñozsabater2019),
daily and monthly resolution (CHELSA data, @karger2021). Some variables
required for some models such as relative humidity or vapor pressure
deficit had to be calculated from these variables. The final strategy on
how to use the climate data was different in each model based on the
approach the better fitted the model (Table \@ref(tab:tableModels)). The
soil quality data was provided as a continuous value between 1 and 5
(Soilgrids dataset, @hengl2017, @poggio2021). The protocol also provided
the elevation, slope and aspect for each of the 200 sites
[@asterscienceteam2019], no other spatial information such as
coordinates were given with the exception of iLand that used the blurred
coordinates to obtain detailed soil information values for the different
sites.

The simulations had to be run in the absence of natural disturbances,
one the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
The simulations had to be set up in a way that they will obtain samples
on species specific regeneration rates per decade and per ha. The
modeling teams decided on the simulated area and how they derive these
samples. Further details on how each modeling team prepared the
simulations and the outputs are available in the supplementary material.

The simulations were run from bare ground in the absence of management
to a simulated equilibrium ("potential natural vegetation") with the
current climate as the ultimate goal was to derive the regeneration
niche. The regeneration niche is defined as the passing of a diameter
threshold of 7 and 10 cm (ingrowth). Both the width of the regeneration
niche (i.e., in environmental space) as well as the intensity of the
recruitment process (i.e., the number of ingrowth trees per area and per
unit of time) were of interest. The exact length of the simulation was
decided by the modeling teams (Table \@ref(tab:tableModels)).

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of the following species or
genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
Quercus spp., Alnus glutinosa, Pinus sylvestris.* Three models included
additional species in the simulations (ForClim 1, ForClim 11 and
TreeMig) and three could not simulate all the species (4C, xComp,
TreeMig, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **FORCLIM**
PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED MORE SPECIES.
HERE **TREEMIG** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED
MORE SPECIES AND MISSED ONE. HERE **4C** PROVIDES [ONE
SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE **XCOMP**
PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. The
model aDGVM2 did not have species level option and provided the default
approach with HERE **aDVGM2** PROVIDES [ONE SENTENCE]{.underline} WITH
THE INFORMATION ON THE DEFAULT SET UP FOR THE SPECIES.

Each of the models reported the number of recruits over a 10 year
interval per species and per ha for each of the sites. To better
understand the variation per site each modeling team decided on how to
sample 200 multiple values per site to show the variation in recruitment
for spatial units of 1 ha during the simulated phase of potential
natural vegetation but without providing average of multiple 1-ha
samples. This was done with different strategies such as sampling
simulated data from the same 1-ha plot in the equilibrium over time, or
by sampling several 1-ha plots from the simulated forest at one specific
point in time (in the equilibrium). This resulted in 880000 observation
per model that simulated the 11 species indicated in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
Those models that simulated extra species, the recruitment of those
extra species were aggregated in a category others. The simulated data
per model and the observations used in this study are provided as a
supplementary information / repository.

```{r map, echo=FALSE, fig.cap="The 200 sites with the dominant tree species in the observed recruitment at 10 cm threshold.", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

@@

The simulation results were analyzed in regard of their regeneration
levels, species composition and mortality in regeneration. The observed
data was modeled using a generalized additive model with a negative
binomial distribution to better understand the trend and regeneration
levels of the observed regeneration and the simulation results per
model. The regeneration levels were evaluated against the stand basal
area (the basal area of all the trees in each sample) as a proxy of
light availability for regeneration, the degree-days as a proxy for
temperature and the water balance as a proxy for soil moisture. The mean
recruited values were calculated as the mean across samples per site and
when the regeneration levels were evaluated across the total basal area
and climate gradients the sites.

We evaluated the species composition across models and in relation to
the observed data by calculating the Shannon index that defines the
relative frequency of species referring to basal area where the higher
the value, the higher the diversity of species in a particular site and
sample (Equation \ref{eqpiShannon}).

```{=tex}
\begin{equation}
H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{n,i} = \frac{r_{BA_{n,i}}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where, $s$ is the total number of species simulated $n$ (11 for the
models that simulated all the species and a varied number for those
models that did not consider all the species, see Table
\@ref(tab:tableModels) for more details in the simulated species);
$p_{i}$ is the proportion of the species $i$ calculated as the recruited
basal area ($r_{BA_{n,i}}$) for that species $i$ over the stand basal
area ($Total_{BA_{n}}$) of the sample $n$.

>>>>

---
title: Drivers of tree regeneration remain challenging to be captured by models of
  forest dynamics
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Mats Mahnken,
- Samuel Fischer,
- Xavier Morin,
- Manfred Lexer,
- Christian Hochauer,
- Katarina Merganicova,
- Tobias Mette,
- Werner Rammer,
- Kristin H. Braziunas,
- Paola Mairota,
- Josef Bruna,
- Heike Lischke,
- Daniel Scherrer,
- Simon Scheiter,
- Thomas Hickler,
- Tim Anders,
- Jessica Hetzer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
  bookdown::word_document2: default
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Tree regeneration is a key process in forest dynamics, particularly in
the context of forest resilience after disturbance and climate change.
Models of forest dynamics have been in use since more than 50 years, but
an assessment of their capacity to properly represent regeneration is
still lacking. Using 15 models and model variants that were build to
capture long term (decadal to centennial) forest dynamics at the stand,
landscape and global levels, we simulate tree regeneration across at 200
sites representing environmental gradients across Europe. The simulation
results are benchmarked against data from forest reserves collected at
these 200 sites. We aim to assess how well the models of forest dynamics
are capturing the overall abundance, species composition and mortality
in tree recruitment. The results show that most of the models tend to
over-estimate recruitment rates although this is not always connected to
large mortality rates in the early stages of population dynamics. In the
models, the effect of light availability on recruitment rates is
captured better than the effect of temperature and soil moisture, but
the patterns are not consistent across models. Species richness matches
quite well compared to the ranges observed in the data. Models
simulating higher species diversity at the stand level do not feature
higher diversity in recruitment. *Abies alba* was well captured across
temperature and soil moisture gradients by most models, whereas *Fagus
sylvatica* was generally underestimated. *Picea abies* recruitment was
generally over-estimated, particularly at colder sites. HERE TWO
SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE

@@

Forests provide a wide range of products and services of vital
importance. Under the pressure of climate change and increasing
disturbances, it is becoming ever more important to understand how
forests structure, composition and functions will change, and evaluate
forest capacity to adapt or be resilient after disturbance
[@lindner2010]. A wide range of models of forest dynamics were developed
over the past decades considering the impacts of climate on forest
dynamics [@bugmann2022]. From these studies, it is becoming increasingly
evident that while we have a reasonably good understanding how to model
tree growth [@vanclay1997; @bugmann1996], and substantial efforts have
been dedicated to improving the depiction of tree mortality processes
[@bugmann2019], tree regeneration remains a much less studied process
that is often represented rather coarsely in these models, which is a
clear research gap in the context of climate-induced forest disturbances
and forest resilience.

Tree regeneration depends on multiple processes such as reproduction,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001]. All these processes are all difficult
to assess, and some of them are scarcely understood and thus appear
highly stochastic [@bogdziewicz2021]. Identifying the appropiate level
of complexity for the formulation of the key factors that are leading to
successful tree recruitment is challenging. By tree recruitment, we
refer to the passing of trees to selected diameter threshold (ingrowth)
as a definition of successful regeneration [@käber2022]. Overall, models
are needed to (1) synthesize existing empirical data and explore their
consequences, and (2) assess future tree recruitment e.g. in the context
of global change scenarios. Currently, regeneration processes in dynamic
forest models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022], from entirely ignoring tree regeneration (as done in
classical forest growth models, e.g. @pretzsch2002), across the use of a
few simple environmental filters, as done in most forest "gap" models
(e.g., @shuman2011) and Dynamic Global Vegetation Models (e.g.,
@smith2001; @hickler2012), to complex approaches that incorporate local
feedback from the canopy, multiple ecological processes and often also
small time steps (e.g., @wehrli2006; @seidl2012), or field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time (e.g., @ribbens1994).

Given the current strategies that are used in models of forest dynamics
to represent tree recruitment, model behavior is facing several problems
and that can have strong impacts on forest dynamic, such as very high
levels of recruitment, which then necessitate excess mortality to early
stages to arrive at correct stand structure and composition. Also, often
inadequately high species diversity in regeneration is simulated. This
is characteristic of \'classical\' forest gap models [@gutiérrez2016],
at least as long as the simulations comprise a multitude of species.
Other models use calibration against local canopy-level data to define
appropriate recruitment levels, which is likely to strongly hamper the
general applicability of these models e.g. under scenarios of climate
change.

Correctly capturing the species composition in recruitment as a function
of the presence of seed trees in the canopy is often a particular
challenge, potentially leading to unrealistic successional drift in the
model, which must be corrected by factors that are hard to parameterize
[@lischke2006]. A related issues is the excessive reduction of species
richness due to positive feedback effects such that eventually just
single-species forest stands remain. This is sometimes handled by the
incorporation of a low level of seed influx of all species at all times
(e.g., @lischke2006 ). However, simulated species composition is usually
exceedingly sensitive to that level, whereas the parameters of such
functions are poorly constrained by field data. Furthermore, there is
also often a problem with the insufficient data-based constraint on
parameter values for models that start from very small tree sizes (e.g.,
10 cm of height or even seeds) and track tree development in a
process-oriented manner by considering a multitude of ecological
influences, rather than being statistically driven and emphasizing
recruitment into a larger size class (e.g., @wehrli2006 ).

Thus, a focus on the modeling of tree regeneration processes is sorely
needed and long overdue (cf. @price2001) if we are to make reliable
projections of future forest dynamics, i.e. when the models need to be
operated in extrapolation mode, but also from a fundamental ecological
point of view (increasing systems understanding).

Also in the field of ecology, tree recruitment has not received a lot of
attention [@hanburybrown_forest_2022], although many sub-processes have
been studied in detail. Yet, little data is available that covers all
regeneration processes within one species along environmental gradients,
let alone for a vast suite of species. Thus, the available data are
fragmented, which constitutes a major problem for model building. On the
one hand, monitoring on permanent sample plots (such as National Forest
Inventories) often measure tree recruitment [@zell_modeling_2019],
however often with a design that captures the regeneration intensity
after the stand initiation phase in productive forests [@hallsby2015].
Data from permanent plots, although available across regions, are also
highly heterogeneous, rendering their utility difficult in a model
context (cf. @käber2021). On the other hand, targeted studies to measure
tree recruitment on experimental sites are often limited in spatial
extent, species studied, or sub-set of regeneration processes that are
targeted [@berdanier2016; @collet2006].

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale multi-species harmonized data set
on tree recruitment [@käber2022]. We included models that are based on a
range of 'philosophies', including models operating at the stand to the
global scale as well as the range of models from empirically derived to
"process-based" (e.g. @fabrika2012; @mette2014; @lexer2001;
@bugmann1996). Because of the large variability in recruitment patterns
in nature and the large amount of factors driving recruitment, including
some that are not included in most models, such as deer browsing, we do
not aim at a detailed statistical evaluation of each model. Instead, we
aim at evaluating the general recruitment patterns and magnitudes
simulated by the models, and benchmark the simulated regeneration niche
of multiple species against empirical data along a wide environmental
gradient (temperature, moisture, and light availability).

More specifically, we aimed to answer the following questions. 1) Are
models of forests dynamics capturing correctly tree recruitment levels
across environmental gradients? 2) Are models of forest dynamics
representing correctly the initial composition of tree populations? And
3), do the models of forest dynamics correctly capture the mortality in
tree in-growth?.

@@

Fifteen models of forest dynamics (including two model featuring
variants) were used to assess tree regeneration. The selected models
simulate forest dynamics at the stand, landscape and global scale (Table
\@ref(tab:tableModels)). The approaches used for model construction and
their origin differ, with most of the models featuring a largely
"process-based" approach, whereas two models based on formulations
derived from the statistical analysis of inventory data (SYBILA and
xComp). Some of the models largely rely on the original approach
underlying forest gap models (ForClim 1, ForClim 11, LandClim, Landis-II
); whereas some include plant ecophysiological processes (FORMIND,
PICUS, 4C, iLand). These different approaches underlying the models also
has strongly influenced the formulation of tree recruitment.

Based on their establishment modules, the models can be categorized as
regeneration or recruitment (or ingrowth) models [@vanclay1994;
@konig_tree_2022] (see Table \@ref(tab:tableModels)). Regeneration
models include processes such as flowering and pollination, seed
production, seed dispersal, germination or seedling establishment that
lead to the simulated number of seedlings or saplings. Recruitment
models, in contrast, introduce a certain number of seedlings or saplings
with certain characteristics such as biomass or diameter, without
considering earlier development processes. We can further distinguish
models that feature a feedback in the regeneration, i.e. where the level
and species composition of regeneration is influenced by the existing
adult tree community via the production of seeds, seedlings or saplings.

The starting point for regeneration in the models also differs, ranging
from seedbank, seed, seedling to sapling (i.e., trees much larger than
10 cm height, often at least 2-3 m tall), which implies that the larger
size threshold, the larger the number of ecological processes that must
be aggregated via parameterizations because they cannot be treated
explicitly. This reduces model complexity but comes at the cost of
blurring process representation.

The overall complexity in the establishment modules also varies
considerably. Following @bugmann2022, we can classified seven models
(iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForCLim and TreeMig) as
having rather high complexity in their tree recruitment modules (mean
complexity across all the recruitment process with value above zero in
Table \@ref(tab:tableModels)). The highest complexity is when defining
processes such as the process used for modeling establishment or how is
frost occurrence taken into account in establishment,

Two models, ForClim and ForCEEPs used two alternative model versions.
ForClim variant 1 [@bugmann1996] is based on an establishment module
that adheres closely to the concept introduced by @botkin1972, whereas
ForClim variant 11 is adopting a rather different approach for
establishment model that considers
space filling along with different \'points of entry\' for the action of
specific
environmental factors @huber2020. These two model variants allow us to
evaluate a more process-based establishment module (variant 11) against
a simpler module (variant 1) while the rest of the model structure
remains the same. Similarly,
the two ForCEEPS variants used here allow us to isolate the importance
of the
canopy feedback via seed trees for the quantity and quality of simulated
regeneration, as one variants includes this feedback whereas the other
one does
not.

@@

Recruitment data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most of the models of
forest dynamics have never been confronted with a data-set covering such
gradients over a large number sites, to evaluate how well they are
capturing recruitment. The recruitment observations used here are from a
novel and unprecedented network of sites in forest reserves that
represent the range of environmental gradients in temperature and
precipitation in Europe compiled in the framework of the EuFoRIa network
[@euforia2019] (Figure \@ref(fig:map)). All these forests have not been
managed for at least 10 years prior to their designation as reserves,
and most of them provide time series of natural forest dynamics over
multiple decades and up to ca. 80 years. The censuses periods range from
37 to 155 years, with most of sites with an average of 14 years. The
plot sized ranged from 0.02 ha to 5.52 ha, the recruited values were
further processed and aggregated to provide recruited values per ha
following @käber2022. The data provide information at the tree level,
thus providing information for sequential comparisons such as
recruitment and death between measurements.

We selected 200 sites from this network as the benchmarking data set for
the simulation exercise. Each of these sites features at least two
consecutive measurements. Regeneration was measured at threshold of 7 or
10 cm of stem diameter at all the sites, except for 35 sites where only
regeneration at the 10 cm diameter threshold was available. The
observations used in this study featured 30.900 recruited trees, and
recruitment values per site and sample ranging from zero to 1246 trees
per ha and a mean of 56 trees per ha recruited. For more details on this
unique dataset and the detailed recruitment information please refer to
@käber2022.

@@

The overarching goal of this experiment was to assess the regeneration
niche as it arises from empirical data against its representation in a
wide range of models of forest dynamics. To this end, each moding group
was provided with a detailed protocol (available in the Supplementary
Material X, henceforth "SM") with instructions how to perform the
simulations, providing input variables on climate and soil conditions,
the specification of the simulation setup and the expected output
variables. Neither site information (except for the data specified
below) nor any data on tree regeneration were provided prior to the
simulation. That is, the models were run in "blind flight" mode.

The input variables were collected and aggregated to be adapted to the
different model demands. The time series of climatic variables from 1981
to 2018 were provided in hourly (Era5-land data, @muñozsabater2019),
daily and monthly resolution (CHELSA data, @karger2021). Some variables
required for some models such as relative humidity or vapor pressure
deficit had to be calculated from these variables. The final strategy on
how to use the climate data was different in each model based on the
approach that best suited the model (Table \@ref(tab:tableModels)). Soil
quality data was provided as a continuous value between 1 and 5
(Soilgrids dataset, @hengl2017, @poggio2021). The protocol also provided
the elevation, slope and aspect for each of the 200 sites
[@asterscienceteam2019], no other spatial information such as
coordinates were provided with the exception of iLand which used the
blurred coordinates to derive highly detailed soil information values
for the 200 sites.

The simulations were run in the absence of natural disturbances, only
the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
The simulations were set up to sample species-specific regeneration
rates per decade and per ha in the equilibrium state of the model,
typically entailing a \"spin-up\" run (as we did not provide any forest
data). The modeling teams decided on the simulated area and how they
derived these samples. The exact length of the simulation was also
decided by the modeling teams (Table \@ref(tab:tableModels)). Further
details on how each modeling team prepared the simulations and the
outputs are available in the SM 2.

The simulations were run in the absence of management to a
simulated equilibrium ("Potential Natural Vegetation") with the current
climate, as the ultimate goal was to derive the regeneration niche. This
entails the assumption that (1) the observations from the forest
reserves reflect no traces of forest management and (2) an equilibrium
between forest dynamics and climate. While the former is likely to be
met at least in many of the EuFoRIA reserves, the latter may be more
debatable. However, in the absence of detailed data on the history of
each plot in the EuFoRIa network, some assumptions had to be made. We
define the regeneration niche as the passing of a diameter threshold of
7 or 10 cm respectively (synonym: ingrowth). Both the width of the
regeneration niche (i.e., in environmental space) as well as the
intensity of the recruitment process (i.e., the number of ingrowth trees
per area and per unit of time) were of interest.

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which ingrowth data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
Quercus spp., Alnus glutinosa, Pinus sylvestris.* Two models included
their standard set of species for the simulations which is much larger
(i.e. ForClim 1, ForClim 11). This also applies to TreeMig, but in this
model *Betula pendula* was not included because HERE **TREEMIG**
PROVIDES THE END OF THIS SENTENCE SAYING WHY B pendula WAS NOT INCLUDED.
In three models, fewer than these elveven species were simulated (4C,
xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **4C** PROVIDES
[ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE
**XCOMP** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS
SPECIES. HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
SIMULATED LESS SPECIES. Finally, the
model aDGVM2 is based on a broad suite of Plant Functional Types rather
than
specific species, and hence this model used its default approach of PFTs
for
the simulations.

Each of the models reported the number of recruits by sampling 200 times
over a 10-year interval for each species and per ha for each of the 200
sites. Both the 200 samples per site as well as its average were
reported. The multiple samples per site were used to better understand
the variation within each site. This was done using different
strategies, depending on the model. The strategies included (1) sampling
simulated data from the same 1-ha plot in the equilibrium over time, or
(2) sampling several 1-ha plots from the simulated forest at one
specific point in time (in the equilibrium). This resulted in 880.000
observations per model that simulated the 11 species included in the
protocol (200 sites, 200 samples per site, 11 species and two diameter
thresholds). For those models that simulated extra species, the
recruitment of the extra species was aggregated into a category
"others". The simulated data per model and the observations used in this
study are available in a repository (REF HERE when available).

Some models did not report all the simulated samples or sites to avoid
unrealistic results. The model 4C set up a threshold of 100 m2/ha to
avoid unrealistic stand basal area values, and therefore not all the
samples and sites were reported, this is the consequence of 4C not being
able to do long term simulations without management PLEASE **4C** VERIFY
THIS IS WHAT YOU DID. LPJ-GUESS had 2% of the samples outputs as
grasslands that were considered in the data as zero stand basal area and
zero recruitment for all the tree species simulated. Landis II also
missed X% of the samples HERE **LANDIS-II** PROVIDES [ONE
SENTENCE]{.underline} WHY THEY DID NOT PROVIDE ALL THE SAMPLES.

```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network that were used in the present study, showing the dominant tree species per site in the observed recruitment at the 10 cm threshold.", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

@@

The simulation results were analyzed in regarding the recruitment
levels(i.e., number of recruits per unit time and space), species
composition of the recruitment, and mortality in the recruitment. The
observed data were modeled using a Generalized Additive Model with a
negative binomial distribution to better understand the trend and
regeneration levels of the observed regeneration, relative to the
simulation results per model.

Regeneration levels were evaluated along three key environmental
gradients, i.e. stand basal area (the basal area of all the trees in
each sample) as a proxy of light availability at the forest floor, the
annual degree-day sum (DD) [@allen1976; @fischlin1995] as a proxy for
growing season warmth, and the climatic water balance (CWB) as a proxy
for soil moisture [@speich2019]. The mean values of recruitment were
calculated as the mean across the 200 samples per site.

We evaluated the species composition across models and in relation to
the observed data by calculating the Shannon index based on the relative
frequency of the species in terms of basal area. The higher the value of
the index, the higher the species diversity at a particular site and
sample (Equation \ref{eqpiShannon}).

```{=tex}
\begin{equation}
H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{n,i} = \frac{r_{BA_{n,i}}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where, $s$ is the total number of species simulated that have recruited
basal area above zero in sample $n$; $p_{i}$ is the proportion of the
species $i$ calculated as the recruited basal area ($r_{BA_{n,i}}$) for
that species $i$ over the stand basal area ($Total_{BA_{n}}$) of the
sample $n$.
